% !TEX root = ../main.tex
%

\begin{abstract}
One of the most significant challenges in social science research is the high cost associated with experiments involving human participants. While some studies have explored the use of large language models (LLMs) as imperfect substitutes for human participants, there has been little work on how to define, design, document, and evaluate such experiments. In this paper, we introduce Synthetic Discussion Generation (SDG), a novel NLP task aimed at creating simulated discussions that enable cost-effective pilot experiments. Drawing on existing SDG systems and interdisciplinary literature, we develop a theoretical, task-agnostic framework for designing, evaluating and implementing these simulations.
We argue that the use of proprietary models such as GPT-5 for such experiments is often unjustified in terms of both cost and capability. Our experiments using smaller, quantized models (7B–8B) demonstrate that these models can produce effective simulations at a cost that is more than twenty times lower. By applying this framework to the task of LLM facilitation, we show that synthetic simulations can yield generalizable results by revealing a critical limitation in the capabilities of LLM facilitators: they are unable to determine when to intervene in a discussion, leading to derailment patterns similar to those observed in human interactions. Additionally, we find that different facilitation strategies significantly influence conversational dynamics. Beyond our theoretical SDG framework, we also present a cost-comparison methodology for experimental design, an exploration of available models and algorithms, an open-source Python framework, and a large, publicly available dataset of LLM-generated discussions across multiple models. This dataset provides a valuable resource for fine-tuning facilitators and analyzing the behavior of out-of-the-box LLMs in synthetic discussion settings. Overall, our work demonstrates that experiments using LLM participants can be made significantly more scalable than current approaches, and can produce meaningful insights when rigorous evaluation is applied. 
\end{abstract}