% !TEX root = ../main.tex
%
\section{Related Work}

\subsection{Generative Agent-Based Modeling}
\label{sec:related:discussions}

\acf{sdg} is a domain-specific specialization of \acf{gabm}; which is the use of generative agents (in this case LLMs) for synthetic simulations of believable human behavior~\cite{park2023game}. After the introduction of initial rules-based (and later \ac{ml}-based) models~\cite{epstein1999_abms} they have succeeded at discovering and replicating emergent social behaviors in a variety of simulated scenarios~\cite{Schelling1971,reynold_boids}.
%TODO: complete this
LLMs are able to function as actors in a controlled, simulated environment, much like their rule-based predecessors. Unlike them however, LLMs are able to incorporate some degree of bias and irrationality to their behavior, which bypasses an inherent limitation of prior models~\cite{adornetto2025_gabms}.
This study attempts to adapt current techniques, conclusions and knowledge from the field of \ac{gabm} to the field of \ac{sdg}, which offers its own goals, challenges and limitations.

\ac{sdg} systems include synthetic clones of Reddit \cite{park_simulacra}, Twitter/X \cite{mou_2024}, generic social media sites \cite{tornberg_2023,y_social}, games \cite{Park2023GenerativeAI}, and social experiments \cite{zhou_2024_sotopia}. One approach~\cite{balog_2024} extracts topics and comments from online human discussions and prompt an LLM to continue them. However, it does not use LLM-based user agents to simulate conversational dynamics, nor does it include facilitator agents. Another approach~\cite{ulmer2024} creates synthetic discussions between two roles: an agent controlling a fictional environment and another client-agent interacting with it. These discussions are filtered and used to finetune the agent LLM for a specific task. Finally, a third approach  \cite{abdelnabi_negotiations} generates synthetic negotiations involving multiple agents with different agendas and responsibilities. Our work can be seen as a domain shift of this approach---from negotiation to discussion facilitation---where various user types (e.g., normal users, trolls) engage in discussion overseen by an actor with veto power (in our case, the facilitator).


\subsection{Substituting human subjects for LLMs}
\label{ssec:related:human-llm}

Recent work \cite{grossman_2023, tornberg_2023,argyle2023} argue that synthetic agents have the potential to eventually replace human participants. Indeed, LLMs have demonstrated complex, emergent social behaviors. In a simulated setting, LLM actors have been observed diffusing information through social interaction as well as planning and organizing spontaneous social events without human interaction \cite{Park2023GenerativeAI}. In settings where LLM actors represent distinct factions with goals and agendas, they were shown to execute basic social strategies and adapt their plans in response to the other LLM actors \cite{abdelnabi_negotiations}. Additionally, substituting sociodemographic backgrounds with information from personal interviews enables LLMs to reliably predict further survey responses from interviewed humans~\cite{park2024generativeagentsimulations1000}. While this finding indicates that LLMs can extrapolate data points relating to human preferences, this approach is not usable in experimental settings at scale, since it relies on extensive human interviews and computational resources.

However, significant limitations of LLMs remain in the context of Social Science experiments. Issues include undetectable behavioral hallucinations \cite{rossi_2024}; sociodemographic, statistical and political biases \cite{anthis_2025,hewitt2024predicting,rossi_2024,Taubenfeld2024SystematicBI}; unreliable annotations \cite{jansen_2023,bisbee_2023,neumann_2025,Gligoric2024CanUL}; non-deterministic outputs \cite{atil_2025,bisbee_2023}; and excessive agreeableness \cite{Park2023GenerativeAI,anthis_2025,rossi_2024}. Even if LLMs can mimic some human behaviors, they are ultimately at most a statistical estimate of the ``average'' human encountered in their training data; and thus we can not claim that any such behavior can be indicative of human behavioral patterns without extensive replication studies~\cite{rossi_2024} or system evaluation~\cite{adornetto2025_gabms}.


\subsection{Evaluation}
\label{sec:related:quality}


\subsubsection{Synthetic data quality}
\label{sec:related:quality:diversity}

When creating \acp{sdg} we need to evaluate \emph{synthetic data quality}, as discussions without human involvement tend to deteriorate quickly, often becoming repetitive and low-quality \cite{ulmer2024}. Despite the need for such metrics, methods for quantifying the quality of synthetic data remain limited~\cite{ulmer2024}. 

One approach~\cite{balog_2024} uses a mix of graph-based, methodology-specific, and lexical similarity metrics, many of which depend on human discussion datasets for comparison. The authors also propose a loosely defined “coherence” score, which is LLM-annotated without theoretical grounding. Another approach assesses quality through post-discussion surveys with humans, and by measuring lexical diversity to approximate the variety of opinions expressed~\cite{kim_et_al_chatbot}. 

\emph{``Diversity''}~\cite{ulmer2024} is a discussion-level metric that penalizes repeated text sequences between comments using average pairwise ROUGE-L \cite{lin-2004-rouge} scores. Low diversity points to pathological problems (e.g., LLMs repeating previous comments). On the other hand, extremely high diversity may point to a lack of interaction between participants; a discussion in which participants engage with each other will feature some lexical overlap (e.g., common terms, paraphrasing points of other participants). Specifically, diversity is defined as:

\small
\begin{equation}
	\label{eq:variety}
	\textit{diversity}(d) = 1 - \frac{2}{N_d(N_d-1)}
	\sum_{i=1}^{N_d-1} \sum_{\substack{j=i+1}}^{N_d} ROUGE-L(c(i,d), c(j,d))
\end{equation}
\normalsize
\noindent where $N_d$ the length (in comments) of discussion $d$. While the metric suffers from the limitations of ROUGE scores (mainly the use of exact-word matching), it is computationally efficient, explainable, and independent of any specific domain, topic, and dataset. We will be employing this metric in our analysis, although we note that more robust synthetic quality metrics are needed.

While this metric is sufficient for capturing variety in synthetic discussions, it is not enough to prove that the system is useful for research purposes. To prove this we need to establish that the dynamics observed in the system are realistic, or at least believable~\cite{adornetto2025_gabms}. We discuss which evaluations methods are available to us for \ac{sdg} in the next section.


\subsubsection{GABM evaluation}
\label{sec:related:quality:gabm}

There is currently no consensus on how a \ac{gabm} system should be evaluated~\cite{adornetto2025_gabms}. We can however use a subset of foundational methods from non-generative agent-based models~\cite{collins2024}. We can also use ``face validation''~\cite{rand2011_rigor}, which employs humans to judge observed behaviors and dynamics. Since the original procedures were designed with \ac{ml} and rules-based models in mind, \cite{adornetto2025_gabms} propose an adapted set of guidelines for generative agents, which we briefly outline below.


\paragraph{Foundational and Empirical Methods}
Foundational methods ensure data integrity and consistency with established benchmarks, without requiring manual evaluation:

\begin{itemize}[nosep, noitemsep]
	\item \textbf{Data analytics}: A prerequisite for all further analysis. Includes rigorous cleaning, organization, and statistical analysis to ensure data integrity.
	
	\item \textbf{Docking}: Also known as model-to-model comparison, this involves comparing a simulation's outputs with those of other state-of-the-art ABMs to verify alignment. It is particularly useful when a recognized reference model exists.
	
	\item \textbf{Empirical validation}: This strategy uses real-world historical or observational data to ensure that model predictions are grounded in reality. It is the primary means of evaluating a model's accuracy in replicating known phenomena.
	
	\item \textbf{Sampling}: Sampling involves exploring the simulation space using various techniques to understand how changes in input variables (e.g., decision thresholds) impact outcomes. This often requires extensive computational resources due to repeated simulation runs, and is therefore non-applicable for systems with non-\ac{ml} and non-rules-based agents.
	
	\item \textbf{Visualization}: This method utilizes graphical representations to analyze emergent behaviors and patterns. While powerful for detecting anomalies, it is subject to qualitative interpretation biases.
\end{itemize}


\paragraph{Face-Validation Methods}
Face-validation focuses on whether model processes and outcomes appear ``reasonable'' to experts, and is especially critical for \acp{gabm} where natural language and social dynamics are central. Unlike empirical methods, these do not strictly require external datasets, but do require manual human labor.

\begin{itemize}[nosep, noitemsep]
	\item \textbf{Survey-based validation}: Human evaluators---often numbering in the dozens or hundreds---are engaged to rank or score agent behaviors based on their perceived believability and social intelligence. Believability in this case refers to the agents' \emph{behaviors}; in the case of LLMs it does not refer to human-like text, which should be taken for granted given the primary goal of LLMs being human text replication~\cite{rossi_2024}. \cite{adornetto2025_gabms} suggest that the agents themselves could be ``interviewed'', and their responses be judged by humans.
	
	\item \textbf{Observational studies}: These analyze human behavior in real-world settings to compare them with generative agents. A notable example is the S3 (Social-network simulation system), which used LLM-agents to simulate dynamics in online social networks. The authors validated that the agents successfully replicated human emotional shifts and information diffusion patterns in response to real-world triggers, by comparing their responses to human behaviors documented, among others, during a real-life disaster in Japan.
	
	\item \textbf{Expert reviews}: Domain experts conduct structured walkthroughs or ``immersive assessments'' where they may even adopt the perspective of an agent to directly evaluate the coherence and plausibility of interactions. This approach has been used in work relating to the effectiveness of LLM facilitators~\cite{cho-etal-2024-language}, although this was only allowed by the setup used, the relatively low number of samples, and the ability to pay participants relatively low wages.
\end{itemize}


\subsection{Online Facilitation}

\subsubsection{Human Facilitation}

Human facilitators\footnote{Called ``moderators'' in some publications---the term is interchangeable in this case~\cite{korre2025evaluation}.} play a crucial role in online communities, serving as key contributors to maintaining social cohesion, encouraging participation, and guiding discussions in a constructive manner~\cite{bcause2023}. Their role is often informal and community-driven, with facilitators typically selected by the community members themselves~\cite{bcause2023} and functioning as de-facto community leaders~\cite{seering_self_moderation}. As such, they are not only moderators in the traditional sense but also volunteers who embody the values and norms of the group~\cite{korre2025evaluation}. The diversity of approaches to human moderation reflects the complexity of online interaction, with various strategies employed to manage discourse, resolve conflicts, and ensure equal participation~\cite{gao-etal-2025-moderation,schroeder-etal-2024-fora,chen-etal-2025-whow, park_et_al_2012_facilitation}. 


\subsubsection{LLM Facilitators}
Unlike classification models traditionally used in online platforms, LLMs can actively facilitate discussions \cite{korre2025evaluation}. 
They can warn users for rule violations \cite{Kumar_AbuHashem_Durumeric_2024}, monitor engagement \cite{schroeder-etal-2024-fora}, aggregate diverse opinions \cite{small-polis-llm}, and provide translations and writing tips---which is especially useful for marginalized groups \cite{Tsai2024Generative}. These capabilities suggest that LLMs may be able to assist or even replace human facilitators in many tasks \cite{small-polis-llm,korre2025evaluation}.

Simple rule-based models can enhance discussions~\cite{kim_et_al_chatbot}, although their approach was largely confined to organizing according to a specific social science (not facilitative) strategy, and balancing user activity. LLM facilitators have been used in excerpts of human discussions with a single participant~\cite{cho-etal-2024-language}. The models provided high-quality feedback to users, although they struggle to make them more respectful and cooperative. In contrast to both works, ours uses exclusively LLM participants and LLM facilitators, and tests the latter in an explicitly toxic and challenging environment.