% !TEX root = ../main.tex
%
\section{Discussion}


\paragraph{Conclusion}
While recent studies predominantly rely on proprietary LLMs for simulations, we argue that this choice is both excessive in terms of the required capabilities and unscalable due to high inference costs. To support this position, we demonstrated that experiments using small open-source models can reduce inference costs by more than twenty times, while still preserving emergent social properties and producing relatively convincing discussions. However, downscaling the models required the development of a unified, task-agnostic theoretical framework for \ac{sdg}, in which we identified key components and requirements, and proposed an evaluation schema to validate these elements.

Despite ongoing skepticism regarding the use and generalizability of LLM simulations, we showed that \ac{sdg} systems can yield meaningful insights by uncovering a critical limitation of LLM facilitators: their tendency to constantly intervene in discussions. Furthermore, we demonstrated that different facilitation strategies significantly influence the patterns of facilitation observed. Overall, our work represents a methodological shift in the design and execution of \ac{sdg} experiments. Our contributions include:
(1) a theoretical framework for designing, documenting, and evaluating \ac{sdg} systems,
(2) an open-source Python framework implementing this methodology,
(3) a cost-comparison methodology for social science experiments,
(4) an exploration of available models and algorithms, and
(5) a large-scale dataset of synthetic discussions. 


\paragraph{Limitations}
Our research is fundamentally restricted by the unavailability of robust evaluation metrics for synthetic text data; while the diversity metric used extensively in this study is a step towards the right direction, its utility remains limited due to exact-word matching and well-known issues with ROUGE. The field of \ac{gabm} is also lacking in robust evaluation strategies not involving extensive, specialized and costly human annotation. Our simulations are also fundamentally constrained by the stereotypical and repetitive speech exhibited by many instruction-tuned models.


\paragraph{Future work}
Future work should thus focus on further ways to robustly evaluate synthetic discussions and replication of social behaviors by LLMs. Additionally, we believe that \ac{sdg} is a promising direction which can be applied to many domains outside of LLM facilitation. Qualitative extensions of this work could involve the use of a dimorphic setup to differentiate participants and facilitators as discussed in \S\ref{sec:scalability:experimental}, the use of non-finetuned models for more ``authentic'' generation (\S\ref{sec:scalability:model}), and sophisticated turn-taking algorithms that give agency to the LLM participants (\S\ref{sec:methodology:management:turn-taking}). Finally, while our research has highlighted that changing facilitation strategies fundamentally changes the behavior of the LLM facilitator, the exact changes should be studied through replication studies with humans.


\paragraph{Ethical considerations}
Synthetic discussions involving LLMs could be exploited by malicious actors to train them at performing unethical tasks \cite{majumdar_2024_nefarious,MARULLI20245340,li_2025_vulnerable}. Additionally, our research includes elements of LLM jail-breaking using conversational context, which could be used to poison discussions using safety-aligned models, and ongoing research is addressing these vulnerabilities \cite{wang_2025_risk}. Furthermore,the use of LLMs inherently risks skewing moderation systems towards the predominant demographics best represented in their training data. \ac{sdb} prompts are a necessary but insufficient step towards avoiding this \cite{rossi_2024,anthis_2025,burton2024large}. 

\paragraph{AI use statement}
We have used small, open source LLMs (primarily LLaMa-8b and Qwen-7B) to help in small style changes throughout this document. We also used them in conjunction with a GPT-5 model \cite{openai2024gpt4technicalreport} to write initial documentation for our framework, as well as to generate some of the code for the plots presented in this paper. 